---
title: "class8"
author: "Luan Tran"
date: "4/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-means clustering

Lets start with an example of running the **kmeans()** function

```{r}
# Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
?kmeans
```

Use the kmeans() function setting k to 2 and nstart=20
Inspect/print the results
Q. How many points are in each cluster?
Q. What ‘component’ of your result object details
 - cluster size?
 - cluster assignment/membership?
 - cluster center?
Plot x colored by the kmeans cluster assignment and
 add cluster centers as blue points
 
```{r}
km <- kmeans(x, centers = 2, nstart = 20)
km
```

```{r}
km$size
```
```{r}
km$cluster
```

```{r}
plot(x, col=km$cluster)

```

 ## Hierarchical Clustering Example
 
 We must give the **hclust()** function a distance matrix not the raw data as input
 
```{r}
# Distance matric calculation
d <- dist(x)
#clustering
hc <- hclust(dist(x))
plot(hc)
abline(h=2, col="red")
abline(h=2.8, col ="blue")
```
 
 
```{r}
cutree(hc, k=2)
```
```{r}
cutree(hc, k=3)
```
 
```{r}
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
```
 
```{r}
plot(x)
```

```{r}
col <- as.factor( rep(c("c1","c2","c3"), each=50) )

plot(x, col=col)
cutree(x, k = 3)
```
 
 
```{r}
plot(x, col=col$cutree)
```
 
 # PCA: Principal Component Analysis
 
 We will use the **prcomp()** function for PCA
 
```{r}
mydata <- read.csv("https://tinyurl.com/expression-CSV",
 row.names=1) 
```
 
 
```{r}
colnames(mydata)
```
 
```{r}
ncol(mydata)
```
 
Run our PCA analysis on the transpose of our data
```{r}
pca <- prcomp( t(mydata), scale = TRUE)
```
 
 
```{r}
attributes(pca)
```
 
 ## A basic PC1 vs PC2 2-D plot
```{r}
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2")
```
 
 ## Percent variance is often more information to look at 
 Calculate the percent variance captured in each PC
 
```{r}
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1) 


head(pca.var.per)
```


## A vector of colors for wt and ko samples
```{r}
colvec <- colnames(mydata)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
 xlab=paste0("PC1 (", pca.var.per[1], "%)"),
 ylab=paste0("PC2 (", pca.var.per[2], "%)")) 
```
 
 